Inference of proto-neutron star properties from gravitational-wave data in core-collapse supernovae

### 2020-11-20 simulations.tex
cleanup biblio

Unlike other methods used in GW astronomy the algorithm proposed in {this work} does not require accurate
waveforms {in order to infer the properties of the PNS.} Instead, it relies on the evolution of the oscillation
frequency of some particular modes visible in the GW spectrum.
The frequency of these modes depends in a universal way on the surface gravity of the PNS, $r\equiv M_{\rm PNS}/R_{\rm PNS}^2$ \citep{Torres:2019b}. Therefore, if at a given time GW emission is observed at a certain
frequency $f$ then the value of the surface gravity can be determined unequivocally, within a certain error,
regardless of the details of the numerical simulation.

@@

%{We have used two different numerical codes in our numerical simulations.} 
%CoCoNuT
%(one-dimensional models) and AENUS-ALCAR
%\citep{Just_et_al__2015__mnras__Anewmultidimensionalenergy-dependenttwo-momenttransportcodeforneutrino-hydrodynamics}
%(one- and two-dimensional models). 
%CoCoNuT
%\citep{Dimmelmeier:2002,Dimmelmeier:2005} is a code for general
%relativistic hydrodynamics coupled to the Fast Multigroup Transport
%scheme \citep{Mueller_Janka_2015_FMT} providing an approximate
%description of the emission and transport of neutrinos.
Both the {\it model set} and {\it test set} simulations have been generated using the AENUS-ALCAR code
\citep{Just_et_al__2015__mnras__Anewmultidimensionalenergy-dependenttwo-momenttransportcodeforneutrino-hydrodynamics}
which combines special relativistic (magneto-)hydrodynamics, a modified
Newtonian gravitational potential approximating the effects of general
relativity \citep{Marek_etal__2006__AA__TOV-potential}, and a spectral
two-moment neutrino transport solver
\citep{Just_et_al__2015__mnras__Anewmultidimensionalenergy-dependenttwo-momenttransportcodeforneutrino-hydrodynamics}.
All siimulations include the relevant reactions between matter and neutrinos of all
flavours, i.e., emission and absorption by nucleons and nuclei,
electron-positron pair annihilation, nucleonic bremsstrahlung, and
scattering off nucleons, nuclei, and electrons.

For the {\it model set}, we use the \pcd{$18$ }spherically symmetric (1D) simulations of \citep{Torres:2019a}
including progenitors with zero-age main sequence (ZAMS) masses in the range 
$M_{\mathrm{ZAMS}}=$ \unit[11.2 -- 75]{$\Msol$}. The set contains simulations using  
six different EOS. Details can be found in
 \citep{Torres:2019a}. The reason to use 1D simulations for the model set
 is that their computational cost is significantly smaller than that of multidimensional
 simulations which allows us to accumulate the statistics necessary to build a good model for $r(f)$.
 {For each time of each simulation we compute the ratio $r$ and the frequency of the $^2g_2$ mode by means of the linear analysis
 described in \cite{Torres:2018,Torres:2019a,Torres:2019b}. }

@@

{For the {\it test set}, we use $8$ axisymmetric (2D) simulations performed with the {\sc AENUS-ALCAR} code
(see Table~\ref{Tab:2dSimList} for a list of models).
All of these simulations but model \texttt{s20S} use a selection of progenitors with masses in the range} $M_{\mathrm{ZAMS}}=$ \unit[11.2 -- 40]{$\Msol$}
 evolved through the hydrostatic phases by
\cite{Woosley_Heger_Weaver__2002__ReviewsofModernPhysics__The_evolution_and_explosion_of_massive_stars}.
We performed one simulation of each stellar model using the EOS of \cite{Lattimer_Swesty__1991__NuclearPhysicsA__LS-EOS} with
an incompressibility of $K=$ \unit[220]{MeV} (LS220) and added
comparison simulations with the SFHo EOS
\cite{Steiner_et_al__2013__apj__Core-collapseSupernovaEquationsofStateBasedonNeutronStarObservations}
and the GShen EOS 
\cite{Shen_et_al__2011__prc__Newequationofstateforastrophysicalsimulations}
for the {progenitor} with $M_{\mathrm{ZAMS}}=$ \unit[15]{$\Msol$}. To this set of
simulations we add the waveform of a 2D model used in
\cite{Torres:2019a}, denoted \texttt{s20S}. It corresponds to a star
with the same initial mass, $M_{\mathrm{ZAMS}}=$ \unit[20]{$\Msol$}, as for
one of the other seven axisymmetric simulations, but was taken from a
newer set of stellar-evolution models
\cite{Woosley_Heger__2007__physrep__Nucleosynthesisandremnantsinmassivestarsofsolarmetallicity}.
It was evolved with the SFHo EOS.

>>>>

Unlike other methods used in GW astronomy the algorithm proposed in {this work} does not require accurate
waveforms {in order to infer the properties of the PNS.} Instead, it relies on the evolution of the oscillation
frequency of some particular modes visible in the GW spectrum.
The frequency of these modes depends in a universal way on the surface gravity of the PNS, $r\equiv M_{\rm PNS}/R_{\rm PNS}^2$~\citep{Torres:2019b}. Therefore, if at a given time GW emission is observed at a certain
frequency $f$ then the value of the surface gravity can be determined unequivocally, within a certain error,
regardless of the details of the numerical simulation.

@@

%{We have used two different numerical codes in our numerical simulations.} 
%CoCoNuT
%(one-dimensional models) and AENUS-ALCAR
%\citep{Just_et_al__2015__mnras__Anewmultidimensionalenergy-dependenttwo-momenttransportcodeforneutrino-hydrodynamics}
%(one- and two-dimensional models). 
%CoCoNuT
%\citep{Dimmelmeier:2002,Dimmelmeier:2005} is a code for general
%relativistic hydrodynamics coupled to the Fast Multigroup Transport
%scheme \citep{Mueller_Janka_2015_FMT} providing an approximate
%description of the emission and transport of neutrinos.
Both the {\it model set} and {\it test set} simulations have been generated using the AENUS-ALCAR code~\citep{Just:2015}
which combines special relativistic (magneto-)hydrodynamics, a modified Newtonian gravitational potential approximating the effects of general relativity~\citep{Marek_etal__2006__AA__TOV-potential}, and a spectral
two-moment neutrino transport solver~\citep{Just:2015}.
All siimulations include the relevant reactions between matter and neutrinos of all
flavours, i.e., emission and absorption by nucleons and nuclei,
electron-positron pair annihilation, nucleonic bremsstrahlung, and
scattering off nucleons, nuclei, and electrons.

For the {\it model set}, we use the $18$ spherically symmetric (1D) simulations of~\citep{Torres:2019a} including progenitors with zero-age main sequence (ZAMS) masses in the range $M_{\mathrm{ZAMS}}=$ \unit[11.2 -- 75]{$\Msol$}. The set contains simulations using six different EOS. Details can be found in~\citep{Torres:2019a}. The reason to use 1D simulations for the model set is that their computational cost is significantly smaller than that of multidimensional simulations which allows us to accumulate the statistics necessary to build a good model for $r(f)$.
For each time of each simulation we compute the ratio $r$ and the frequency of the $^2g_2$ mode by means of the linear analysis described in \cite{Torres:2018,Torres:2019a,Torres:2019b}.

@@

{For the {\it test set}, we use $8$ axisymmetric (2D) simulations performed with the {\sc AENUS-ALCAR} code
(see Table~\ref{Tab:2dSimList} for a list of models).
All of these simulations but model \texttt{s20S} use a selection of progenitors with masses in the range} $M_{\mathrm{ZAMS}}=$ \unit[11.2 -- 40]{$\Msol$}
 evolved through the hydrostatic phases by~\citep{Woosley_Heger_Weaver__2002__ReviewsofModernPhysics__The_evolution_and_explosion_of_massive_stars}.
We performed one simulation of each stellar model using the EOS of~\citep{Lattimer_Swesty__1991__NuclearPhysicsA__LS-EOS} with an incompressibility of $K=$ \unit[220]{MeV} (LS220) and added comparison simulations with the SFHo EOS~\citep{Steiner_et_al__2013__apj__Core-collapseSupernovaEquationsofStateBasedonNeutronStarObservations} and the GShen EOS~\citep{Shen_et_al__2011__prc__Newequationofstateforastrophysicalsimulations} for the {progenitor} with $M_{\mathrm{ZAMS}}=$ \unit[15]{$\Msol$}.
To this set of simulations we add the waveform of a 2D model used in~\citep{Torres:2019a}, denoted \texttt{s20S}. It corresponds to a star with the same initial mass, $M_{\mathrm{ZAMS}}=$ \unit[20]{$\Msol$}, as for one of the other seven axisymmetric simulations, but was taken from a newer set of stellar-evolution models~\citep{Woosley_Heger__2007__physrep__Nucleosynthesisandremnantsinmassivestarsofsolarmetallicity}. It was evolved with the SFHo EOS.
